---
title: "homework-04"
format: html
editor: visual
author: izzy tector
execute: 
  warning: false
  message: false
---

## Homework 4

Izzy Tector

5/23/23

```{r setting-up}
library(tidyverse)
library(lterdatasampler)
library(here)
library(extrafont)
library(performance)
library(broom)
library(flextable)
library(ggeffects)
library(car)
library(naniar)
```

# Problem 1

1.  Mathematical:

    Null Hypothesis:

    There is no linear relationship between the trout perch length and weight.

    Alternative Hypothesis:

    There is a linear relationship between the trout perch length and weight.

Biological:

Null: The trout perch length does not have any effect on the weight of the fishes across all years.

Alternative: The trout perch length predicts the weight of the fished across all years.

2.  

```{r loading-data}
fish <- read_csv(
  here("data", "ntl6_v12.csv")
)
```

```{r filtering-data}
trout_perch <- fish %>% 
  filter(spname == "TROUTPERCH")
```

visualizing missing data

```{r missing-data-vis}
gg_miss_var(trout_perch)
```

a\. Since there is a lot of the weight data missing, this could heavily affect the my hypothesis and testing of the relationship between the length and the weight of the trout perches.

3.  

```{r exploratory-data-vis}
ggplot(data = trout_perch, aes(x = length, y = weight)) +
  geom_point()
```

```{r model-object}
modelobject <- lm(weight ~ length, data = trout_perch)

modelobject
```

4.  

```{r linearization-plots}
par(mfrow = c(2, 2))
plot(modelobject)
```

```{r dev-off, results= false}
dev.off()
```

5.  

Residuals vs fitted: The data generally fits the dotted line, showing that there is a linear relationship between the variables.

Normal Q-Q: The data fits the linear line fairly well except for the outliers, showing the data is normally distributed.

Scale-Location: The scale-location plot shows the data has equal variation or homoskedasticity across the data frame because it follows the dotted line.

Residuals vs Leverage: Cook's distance shows us that we have an outlier that is affecting our model from our data.

6.  

```{r}
model_summary <- summary(modelobject)

model_summary
```

(F1,288 = 1281, p \< 0.001, R\^2 = 0.82

7.  

```{r anova}
model_anova <- anova(modelobject)

model_anova
```

```{r}
# don't name this chunk!

model_anova_table <- tidy(model_anova) %>% 
  mutate(across(sumsq:meansq, ~ round(.x, digits = 5))) %>% 
  mutate(statistic = round(statistic, digits = 1)) %>% 
  mutate(p.value = case_when(
    p.value < 0.001 ~ "< 0.001"
  )) %>% 
  mutate(term = case_when(
    term == "length" ~ "Length (mm)",
    TRUE ~ term
  )) %>% 
  flextable() %>% 
  set_header_labels(df = "Degrees of Freedom", 
                    sumsq = "Sum of squares",
                    meansq = "Mean squares",
                    statistic = "F-statistic",
                    p.value = "p-value")


model_anova_table
```

8.  
