---
title: "homework-04"
format: html
editor: visual
author: izzy tector
execute: 
  warning: false
  message: false
---

## Homework 4

Izzy Tector

5/23/23

Link to repo [here](https://github.com/itector/ENVS-193DS_homework-04_tector-izzy.git).

```{r setting-up}
library(tidyverse)
library(lterdatasampler)
library(here)
library(extrafont)
library(performance)
library(broom)
library(flextable)
library(ggeffects)
library(car)
library(naniar)
```

# Problem 1

1.  Mathematical:

    Null Hypothesis:
    Hₒ : β₁ = 0

    There is no linear relationship between the trout perch length and weight.

    Alternative Hypothesis:

    Hₐ : β₁ ≠ 0
    There is a linear relationship between the trout perch length and weight.

Biological:

Null: The trout perch length does not have any effect on the weight of the fishes across all years.

Alternative: The trout perch length predicts the weight of the fished across all years.

2.  

```{r loading-data}
fish <- read_csv(
  here("data", "ntl6_v12.csv")
)
```

```{r filtering-data}
trout_perch <- fish %>% 
  filter(spname == "TROUTPERCH")
```

visualizing missing data

```{r missing-data-vis}
# visualizing the missing data in the data frame using gg_miss_var
gg_miss_var(trout_perch)
```

**Figure 1. Trout perch data frame missing value visualization throughout all years.** Variables of trout perch data collected from North Temperate Lakes LTER are represented on y axis. X axis represents the number of missing data points from each of the variables. Sex, flag, fishpart, sampletype, and weight all have missing data points.

a\. Since there is a lot of the weight data missing, this could heavily affect the my hypothesis and testing of the relationship between the length and the weight of the trout perches.

3.  

```{r exploratory-data-vis}
# create an exploratory data visualization to assess whether the data could have a linear relationship
ggplot(data = trout_perch, aes(x = length, y = weight)) +
  geom_point()
```

```{r model-object}
# run the linear regression model and create an object to create plots out of
modelobject <- lm(weight ~ length, data = trout_perch)

modelobject
```

4.  

```{r linearization-plots}
# create plots of the model object and section it off into a table of 4 plots
par(mfrow = c(2, 2))
plot(modelobject)
```

```{r dev-off}
# document would not render with the resultf = false on
dev.off()
```

5.  

Residuals vs fitted: The data generally fits the dotted line, showing that there is a linear relationship between the variables.

Normal Q-Q: The data fits the linear line fairly well except for the outliers, showing the data is normally distributed.

Scale-Location: The scale-location plot shows the data has equal variation or homoskedasticity across the data frame because it follows the dotted line.

Residuals vs Leverage: Cook's distance shows us that we have an outlier that is affecting our model from our data.

6.  

```{r}
# use summary() to create a summary of the model object data and linear regression model
model_summary <- summary(modelobject)

model_summary
```

(F1,288 = 1281, p \< 0.001, R\^2 = 0.82

7.  

```{r anova}
# run an anova test to see sum of squares and mean squares
model_anova <- anova(modelobject)

model_anova
```

```{r}
# don't name this chunk!

# create a table of the anova test values
model_anova_table <- tidy(model_anova) %>% 
  # clean the data up by rounding the numbers
  mutate(across(sumsq:meansq, ~ round(.x, digits = 5))) %>% 
  mutate(statistic = round(statistic, digits = 1)) %>% 
  mutate(p.value = case_when(
    # clean the p-value up by changing it to be <0.001
    p.value < 0.001 ~ "< 0.001"
  )) %>% 
  # clean up the names of the titles
  mutate(term = case_when(
    term == "length" ~ "Length (mm)",
    TRUE ~ term
  )) %>% 
  # make the table a flex table
  flextable() %>% 
  set_header_labels(df = "Degrees of Freedom", 
                    sumsq = "Sum of squares",
                    meansq = "Mean squares",
                    statistic = "F-statistic",
                    p.value = "p-value")


model_anova_table
```

8.  

Both the ANOVA table and the summary object gives the F-statistic, P value, and degrees of freedom. These values were the same for both with an F value of 1281 (1280.8 for the ANOVA table), a p-value \<0.001 (confidence level of 95%), and the degrees of freedom being 1 and 288 (sample size of 290). The ANOVA table gives the extra information of the sum of squares and the mean squares. and the summary gives residuals values which the ANOVA table does not have.

9.  

The test results have proven that the trout perches length is an effective predictor of the weight of the fishes throughout all years. This was shown through the p-value, given as 2.2e-16, which is \<0.001, so I reject the null hypothesis. The sum of squares (1432) and the mean squares (1432) of the length of the fish show that the variation in fish weight is proportional to the variation in fish length. The R-squared value (0.8164) shows that the data follows a linear regression line. The model equation would be for every increase of one millimeter of fish length, the weight would be predicted to increase by 0.20 +/- 0.006 grams.

10. 

```{r predictions}
# create an object using ggpredict
predictions <- ggpredict(modelobject, terms = "length")

predictions
```

```{r plot}
# create a plot visualizing the data and the predictions
plot_predictions <- ggplot(data = trout_perch, 
                           aes(x = length, y = weight)) +
  # first plot the underlying data from trout_perch
  geom_point() +
  # then plot the predictions
  geom_line(data = predictions, 
            aes(x = x, y = predicted), 
            color = "orchid", linewidth = 1) +
  # then plot the 95% confidence interval from ggpredict
  geom_ribbon(data = predictions, 
              aes(x = x, y = predicted, ymin = conf.low, ymax = conf.high), 
              alpha = 0.2) +
  # theme and meaningful labels
  theme_bw() +
  labs(title = "Trout perch length predicting weight trends throughout all sample years", 
       x = "Length (mm)",
       y = "Weight (g)") + 
  theme(text = element_text(family = "Times New Roman")
  )


plot_predictions
```

**Figure 2. Trout perch length predicting weight trends throughout all sample years**. Length (mm) of trout perch collected from North Temperate Lakes LTER represented on the x axis. Weight (g) is represented on the y axis. Dots representing the data points with a line of best fit (magenta) for linear regression model. Transparent ribbon representing the confidence level of the data.
